---
title: "Linear_Regression"
author: "Matthew Gregory"
date: "Tuesday, September 22, 2015"
output: html_document
---

We are interested in predicting the price of a diamond based on 9 predictors. We want to know if we are getting ripped off for the price we pay for a diamond compared to historical prices.  
First we load the appropriate libraries (makes sure you have them installed).

```{r, echo=FALSE, warning=FALSE}
library("ISLR")
library("ggplot2")
library("GGally")
library("scales")
library("memisc")
library("dplyr")

```

Then we read in the diamonds data and assign it to an object called `mydata`. We inspect the structure of the dataframe using `str`.

```{r, echo=FALSE}
mydata <- diamonds
str(mydata)
```
The dataframe seems to have the factors in the correct order and numerical variables assigned as such.  
```{r, echo = TRUE}
#WARNING takes a 10 minutes to run!
#diasamp <- mydata[sample(1:length(mydata$price), 10000),]
#ggpairs(diasamp, params = c(shape = I("."), outlier.shape = I(".")))
```

What's happening is that ggpairs is plotting each variable against the other in a pretty smart way. In the lower-triangle of the plot matrix, it uses grouped histograms for qualitative-qualitative pairs and scatterplots for quantitative-quantitative pairs.  In the upper-triangle, it plots grouped histograms for qualitative-qualitative pairs (using the x-instead of y-variable as the grouping factor), boxplots for qualitative-quantitative pairs, and provides the correlation for quantitative-quantitative pairs.  

  
Let's define a function to inspect the cube root of the carat, it's volume.

```{r, echo=TRUE}
cubroot_trans = function() trans_new("cubroot", transform= function(x) x^(1/3), inverse = function(x) x^3 )
```
With so many data points overfitting is a concern, let's jitter and make each point more transparent.

```{r}
p1 <- ggplot( data=mydata, aes(carat, price)) +
    geom_point(alpha = 0.5, size = .75, position="jitter") +
    scale_x_continuous(trans=cubroot_trans(), limits = c(0.2,3),
        breaks = c(0.2, 0.5, 1, 2, 3)) +
    scale_y_continuous(trans=log10_trans(), limits = c(350,15000),
        breaks = c(350, 1000, 5000, 10000, 15000)) +
    theme_bw() +
    ggtitle("Price (log10) by Cubed-Root of Carat")
p1
```

Our 2008 data has some problems but let's go ahead and model price using linear regression.

```{r, echo = TRUE}

mydata$logprice <- log(mydata$price)
m1 <- lm(logprice~  I(carat^(1/3)), 
    data=mydata[mydata$price < 10000,])
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut )
m4 <- update(m3, ~ . + color + clarity)
 

mtable(m1, m2, m3, m4)

```
We can use the `summary` function on each of these models to assess how much variation within the data they explain.
  
We can use the model to see if we are being charged too much for a diamond we are interested in.

```{r, echo = TRUE}
# Round 1.00 Very Good I VS1 $5,601
thisDiamond <- data.frame(carat = 1.00, cut = "Very Good", color = "I", clarity="VS1")
modEst <- predict(m4, newdata = thisDiamond, interval="prediction", level = .95)
exp(modEst)
```
The predicted price from the model is lower than what they are asking for by $600 - hussle!